---
title: "Larval Billfish Identifications"
author: "Andrea Schmidt"
date: "2024-06-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#libraries#####
library(tidyverse)
library(lubridate)
library(readxl)
#data read in#####
bf<-read.csv("C:/github/billfish_2024/MASTER_BillfishInventory_WHIP+Slicks_Istiophoridae-Xiphiidae_Counts-Sizes_20241305 - BillfishInventory_WHIP+Slicks_Istiophoridae-Xiphiidae_Counts-Sizes_20240202.csv")
```

## Brief Introduction

Unknown istiophorid larvae were inventoried, their lengths measured and tissue collected. These larval tissues were then extracted using the high throughput HOTSHOT extraction method (citation...UCONN) and amplified using billfish specific primers targeting the cyt b region after Hyde et al 2005. Once tissue extracts were amplifed the post-PCR products were run out on a gel and species were identified based on the size of the amplified DNA. Below is the workflow of how we got from unknown istiphorids to known billfish larvae within a 20 year time series of data collected across Kō Hawaiʻi Pae ʻĀina.

```{data exploration echo=FALSE}
#to pull size data into something useable

```

## Plotting

Larval billfish identifications with proportion of total number of fish of a particular species are as follows:

```{r species by year, echo=FALSE}
ggplot(bf, aes(x=Year,y=taxa, group_by(taxa)))+geom_col()
#trying to sum total number of larvae of a particular taxa across year
mini<-bf %>% group_by(taxa, Year) %>%
  summarize(sum(count))
pie(labels="taxa")
# Package
library(treemap)
treemap(mini,index="taxa",vSize="sum(count)",type="categorical",# Main
           title="",
           palette="Dark2",# Borders:
           border.col=c("black"),
           border.lwds=1,# Labels
           fontsize.labels=0.5,
           fontcolor.labels="white",
           fontface.labels=1,
           bg.labels=c("transparent"),
           align.labels=c("left", "top"),
           overlap.labels=0.5,
           inflate.labels=T)
```

#merging
```{r merge cruise-stations with larval data sheet,  echo=FALSE}
combo<-read.csv("C:/Users/Andrea.Schmidt/Desktop/for offline/combo_whip_slick_short11.csv")
bf2<-bf%>%unite("Site",c(cruise, station),sep= "-")
all <-full_join(bf2, combo, by = join_by(Site)) #join by transect number (not sample)
all2 <-full_join(bf, combo, by = join_by(sample)) #join by transect number (not sample)

```
#Mapping
```{r map of all larvae for all years,  echo=FALSE}
#libraries#####
library(tidyverse)
library(lubridate)
library(readxl)
library(suncalc)
library(measurements)
library(stringr)
library(geosphere)
library(sp) #already in raster
library(raster)# error
library(ggplot2)
library(scales)
library(rgdal)
library(marmap)
library(maps)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
library(viridis)
library(ggplot2)
library(raster)
library(ggnewscale)
library(sf)
#map sampling sites######
world<-ne_countries(scale="medium", returnclass = "sf")
#oahu_raster <- raster(file.path("C:/Users/Andrea.Schmidt/Documents/M&B_larval_dist_csvs/all_hi_bathy.tiff"))
oahu_raster <- raster(file.path("C:/Users/Andrea.Schmidt/Desktop/for offline/billfish_hi_bathy.tiff")) #ETPO_2022(Bedrock) from https://www.ncei.noaa.gov/maps/bathymetry
oahu_df <- fortify(as.bathy(oahu_raster))
str(oahu_df)
oahu_map <- ggplot(data=world) +
  #coord_sf(crs = oahu_raster@crs,xlim=c(-161.5,-154.5),ylim=c(18.85,22.29)) +
  geom_raster(data = oahu_df, aes(x = x, y = y, fill = z)) +labs(fill = "Depth (m)")+
  #scale_fill_gradientn(colours = c(grey(50/100), grey(80/100)),values = rescale(c(min(oahu_df$z, na.rm = TRUE), 0,  0.0001, max(oahu_df$z, na.rm = TRUE))), na.value = "transparent",guide = "none")+
  scale_fill_gradient(high = "lightskyblue1", low = "cornflowerblue",limits=c(-6000,0))+new_scale_fill()+
  theme(panel.background=element_blank(),axis.title.x = element_blank(), axis.title.y = element_blank())+
  theme_bw()+geom_sf()+coord_sf(xlim=c(-153,-161), ylim=c(18.7, 22.2))+
  theme(legend.title=element_text(size=20),legend.text=element_text(size=17),
        legend.direction = "vertical", legend.box = "vertical", axis.text = element_text(size=20))+
  scale_x_continuous(breaks=seq((-161), (-153), 1))+
  scale_y_continuous(breaks = seq(18.7, 22.2,0.5))
oahu_map+geom_point(data=all, aes(y=LAT_DD_start, x=LONG_DD_start, color=taxa))+
  scale_color_viridis_d(option="F")+theme(panel.background=element_blank(),axis.title.x = element_blank(), axis.title.y = element_blank())
```


