---
title: "historic_environmental_data_bf24"
author: "Andrea Schmidt"
date: "2024-07-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(dplyr)
library(oce)
library(gsw)
#Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```

## R Markdown
Environmental data from a variety of platforms, devices and years was assembled for the purposes of creating a dataset of environmental variables associated with each occurrence of larval billfish
and all absences using in-situ oceanographic data collected during cruises from archived/raw
sources (i.e., shipboard thermosalinograph (TSG) and conductivity-temperature-depth (CTD)
casts) for all samples conducted over the time series (temperature, salinity, dissolved oxygen,
and chlorophyll) as well as habitat information (i.e., inside/outside surface slicks) and remotely
sensed ocean products spanning the 20-year period (temperature, salinity, chlorophyll, winds,
sea surface height anomalies, eddy productivity, etc.). First we read in the header files for a given year. IN the following instance, this is 2016:
```{r TSG data from 2016}
#(echo = FALSE)
#focus on TSG data 2004-2006,2017-2018
hdr<-read.delim("C:/github/billfish_2024/TSG-SE-1606/SE-16-06_TSG/SE-16-06_001.hdr")
hdr
ex<-read.delim("C:/github/billfish_2024/TSG-SE-1606/SE-16-06_TSG/SE-16-06_001.XMLCON")
ex
```

## hex file form SBE

This .hex file from 2016 has 319627 rows, each one containing coded data, for example: "B93ECEB87243C210381A787D8F40"

```{r TSG2016 hex file}
pox<-read.delim("C:/github/billfish_2024/TSG-SE-1606/SE-16-06_TSG/SE-16-06_001.hex")
pox
```
so clearly, the next step is to convert this into something useful. 
## Convert hex file into data we can work with
```{r tsg2016 hex to sense}

```
##Notes about oce from youtbe
CTD data is output in .cnv. The function, read.oce() will parse CTD data. oce has the ability to parse non-cnv data types, called oceMagic()
```{open files}
setwd("~/billfish_not_github/HistoricCruiseData_ChrisTokita20190827")
list.files(pattern="*.cnv")
```